### 2023-12-05 manuscript.tex
Made various minor changes in wording for clarity, consistency, or flow

However, no known subsequent studies have determined if similar patterns of conservation are found in the IDRs of other systems. As another foundational model organism with abundant genomic information across many evolutionary lineages~\cite{Yang2018, Miller2018, Kim2021, Gramates2022}, the fruit fly, \textit{Drosophila melanogaster}, is a natural choice for subsequent investigation. Furthermore, given its complex multicellular development process and shared signaling pathways with humans, the findings of such a study would significantly advance our understanding of the role of IDRs in gene regulation as well as human health and disease. The concordance of these results with the previously identified IDR clusters would also have profound implications for the broader mechanisms of IDR evolution. For example, the absence of global patterns of evolutionary signatures across IDRs in \textit{Drosophila} would suggest they are property of IDRs which is unique to yeast. In contrast, the identification of clusters similar to those in yeast would indicate the existence of a taxonomy of IDRs which is conserved across the tree of life. The latter result would represent a significant step towards the creation of resources for the classification of IDRs analogous to those for folded domains such as Pfam~\cite{Mistry2020}, CATH~\cite{Sillitoe2020}, or SCOP~\cite{Andreeva2013, Andreeva2019}.

@@

\section{Results}
\subsection{IDRs are shorter and more divergent than non-disordered regions}
We identified regions with high levels of inferred intrinsic disorder in over 8,500 high quality alignments of single copy orthologs from 33 species in the \textit{Drosophila} genus using the disorder predictor AUCPreD~\cite{Wang2016}. To highlight the unique features of IDR evolution in subsequent analyses, we also extracted a complementary set of regions with low levels of inferred disorder. Both sets were filtered on several criteria, including the lengths of their sequences and their phylogenetic diversity, which yielded 11,445 and 14,927 regions, respectively, from 8,466 unique alignments. In the subsequent discussion, we refer to these sets as the ``disorder'' and ``order'' regions, respectively. To investigate the differences in basic sequence statistics between the two region sets, we first generated histograms from the average length of each region (Fig.~\ref{sfig:regions}). Although both distributions span several orders of magnitude, the order regions are generally longer than the disorder regions, with means of 105 and 245 residues, respectively. We then quantified the sequence divergence in each region by fitting phylogenetic trees to the alignments using amino acid and indel substitution models. The average rates of substitution are significantly larger in the disorder regions, demonstrating that while both sets contain conserved and divergent regions, IDRs are enriched in more rapidly evolving sequences (Fig.~\ref{sfig:regions}). We also searched the database of Pfam domains against the full-length \textit{D. melanogaster} proteins in these alignments and measured their overlap with the disorder and order regions. The results show a clear enrichment in the disorder regions having no or only small amounts of overlap relative to the order regions, indicating that, relative to structured domains, IDRs are especially resistant to homology-based methods of functional annotation (Fig.~\ref{sfig:pfam}).

@@

\subsection{Intrinsic disorder is poorly conserved in some proteins}
Though the substitution models reveal specific patterns of evolution at the level of individual residues, the large amounts of sequence divergence between many orthologous IDRs implies their evolution is not well-described by fine-scale models of residue substitution. Given the growing evidence that IDRs are constrained to conserve distributed properties, we instead turned towards characterizing their evolution in terms of 82 disorder-associated ``molecular features'' obtained from the previous study of IDRs in the yeast proteome. However, before conducting an in-depth analysis of these features, we examined the disorder score traces in greater detail and were struck by the significant variability between species. For each residue in the input sequence, AUCPreD returns a score between 0 and 1 where higher values indicate higher confidence in a prediction of intrinsic disorder. In some alignments, the scores vary by nearly this entire range at a given alignment position even when there is a relatively high level of sequence identity (Fig.~\ref{fig:scores}A).

@@

\subsection{IDRs have three axes of unconstrained variation}
Having calculated the features associated with the sequence segments composing each region in our data set, we then sought to determine if their distributions contained any global structure which would enable us to identify classes with distinct biophysical or functional properties. These distributions are generated by a complex underlying evolutionary process which reflects the combined effects of selection, drift, and mutation. However, to leverage a statistical framework to infer the properties of this process, we fit Brownian motion (BM) models to the features calculated from the segments in each region. BM is a simple model of evolution where continuous traits change through a series of small, undirected steps. Thus, the traits accumulate variation at a constant rate over time but do not on average deviate from their original values. BM models are therefore specified by two parameters: a rate, which describes the speed at which trait variation accumulates, and a root, which describes the ancestral trait value.

We then applied principal components analyses (PCAs) to visualize the major axes of variation of the root and rate parameters for each feature and region. A difficulty with a direct analysis of the parameter estimates, however, is PCAs are sensitive to differences in scaling between variables, and some features have dramatically different intrinsic scales. For example, many compositional features, like fraction\_S, are restricted to the interval [0, 1], whereas SCD is unbounded and can vary from negative to positive infinity. As a result, SCD is responsible for a significant fraction of the overall variance in both parameter distributions (Fig.~\ref{sfig:root_variance}-\ref{sfig:rate_variance}). Therefore, we first normalized the parameters associated with each feature by transforming them into z-scores relative to their proteome-wide distributions.

The first two principal components of the root distributions show little overall structure, though there is a slight enrichment of regions along two axes that correlate with acidic and polar features, respectively (Fig.~\ref{sfig:root}). Likewise, the projections of the rates onto the first two components are largely distributed along the first principal component (Fig.~\ref{fig:rate}A). This observation and the variable amounts of sequence divergence in the regions led us to suspect the first principal component was a measure of the overall rate of sequence evolution. Plotting the first principal component against the sum of the average amino acid and indel rates as measured by substitution models revealed a strong association (Fig.~\ref{fig:rate}B). We then projected the rates along second and third principal components to determine if these higher order components contained any additional structure. The resulting distribution is roughly triangular and contains three major axes of variation, corresponding to rapid changes in the regions' proportions of acidic, glutamine, and glycine residues (Fig.~\ref{fig:rate}C-D). Inspection of regions selected along these axes confirmed the high rates of evolution of these features (Fig.~\ref{fig:rate}E-G). Furthermore, we observe a similar distribution when the rates of the order regions were projected along their second and third principal components, which suggests a lack of constraint along these axes is a general property of rapidly evolving proteins (Fig.~\ref{sfig:order}).

\subsection{A model of constrained evolution can identify signals of conservation}
Though the BM process permits the inference of the rates of feature evolution after accounting for the phylogenetic relationships between species, it does not directly test for their conservation. In fact, under the BM model, trait variation is unconstrained and will increase without bound over time. Instead, evidence of conservation requires comparison to a model where trait variation is constrained. A common choice for modeling the effect of selection on the evolution of a continuous trait is the Ornstein-Uhlenbeck (OU) model. The OU model is similar to the BM model where a trait accumulates variation through a series of small, undirected steps. However, it differs in that the trait is also attracted towards an optimal value where the attraction is proportional to the trait's distance from this value. Under an additional assumption of stationarity that ensures parameter identifiability and estimate consistency, the OU model is therefore specified with three total parameters: the optimal value ($\mu_{OU}$), the fluctuation magnitude ($\sigma^2_{OU}$), and the selection strength ($\alpha$)~\cite{Ho2013, Ho2014}. While the first two parameters are analogous to the root and rate parameters in the BM model, respectively, the selection strength has no equivalent.

@@

We then quantified the type I and II error rates under a hypothesis testing framework. We first calculated the type I error rate as a function of the critical value of the log likelihood ratio for each value of the BM rate separately~(Fig.~\ref{fig:simulation}D). As the overlapping curves indicate the ratios are independent of the true value, we merged the simulations and calculated empirical critical values for 5\% (2.58) and 1\% (4.20) type I error rates~(Fig.~\ref{fig:simulation}E). Using the 1\% critical value, we calculated the type II error rate as a function of the OU parameter values~(Fig.~\ref{fig:simulation}F). The results show a strong dependence on the log ratio of $\alpha$ to $\sigma^2_{OU}$, with the error rate sharply declining to zero when $\log_{10}(\frac{\alpha}{\sigma^2_{OU}}) > 1$. In summary, these results demonstrate that while the parameter estimates of the OU model can have significant bias, when selection is strong relative to the fluctuation magnitude, the likelihood ratio of the two models can reliably signal against a null hypothesis of Brownian motion.

\subsection{Signals of feature conservation are widespread in IDRs}
Having evaluated the statistical properties of the hypothesis testing framework, we next sought to investigate the distribution of significant features across all regions. However, to ensure we would not identify signals of conservation simply as a result of high levels of sequence identity, we restricted our subsequent analyses to disorder regions with a minimum amount of divergence as measured by substitution models (Fig.~\ref{sfig:hierarchy_histogram}). We first quantified the fraction of regions with significant likelihood ratios using the empirical critical values for 1 and 5\% type I error rates, finding that all non-motif features show significance rates that greatly exceed their nominal error rates (Fig.~\ref{fig:conserved}A). The motif features, though, are more variable, with some reaching significance rates similar to those of the non-motif features, whereas many others have rates at or below the nominal values. Generally, the shorter and more degenerate motifs associated with modification sites have higher rates of significance, which may reflect stronger constraints on motifs which are more likely to arise by random mutations. We next calculated the number of significant features in each region using the same critical values. Both yielded highly skewed distributions, indicating the observation of a significant feature increases the probability of observing another significant feature (Fig.~\ref{fig:conserved}B). However, as many features have overlapping definitions, this dependence is expected. To correct for this effect, we therefore randomly shuffled the likelihood ratios between regions for each feature independently and calculated the overall significance rate across all features. For both critical values, the empirical significance rate exceeds the nominal error rate. Thus, constrained feature evolution is widespread in IDRs, with each region containing an average of seven significant features under the 1\% type I error rate.

To illustrate the utility of this method for generating hypotheses to guide the functional dissection of specific proteins, we next compared two regions with elevated values of fraction\_Q and repeat\_QN but different signals of conservation for each (Fig.~\ref{fig:alignment}A-B). Though the alignments of both regions are enriched in glutamine and asparagine residues (Fig.~\ref{fig:alignment}C,E), the first has significant signals of conservation for both features (Fig.~\ref{fig:alignment}D), whereas the second does not (Fig.~\ref{fig:alignment}F). Notably, the first region is from an orthologous group that contains a sequence from the \textit{D. melanogaster} gene {ewg}, which is a transcription factor. As polyglutamine stretches can modulate the activity of transcription factor activation domains, this observation is consistent with the maintenance of an optimal level of transcriptional activation mediated by constraints on glutamine-associated features~\cite{Mitchell1989, Gerber1994}. In contrast, the low likelihood ratios of the second region, whose \textit{D. melanogaster} gene is unannotated, are suggestive of unconstrained glutamine repeat expansion and contraction. Interestingly, this region has highly significant likelihood ratios for several features associated with glycine or phenylalanine repeats despite having few of these residues. However, careful inspection of the alignment reveals several positions where such repeats appear in only one or two sequences, \textit{e.g.} columns 153 and 208. While these strong signals of conservation may represent artifacts of applying a continuous model of trait evolution to features derived from discrete data, they could also indicate true constraints against the presence of flexible or hydrophobic patches in those sequence segments. Taken together, these results highlight this method's ability to identify biochemical features with signals of constrained evolution while also emphasizing the importance of combining those results with biochemical intuition to suggest plausible hypotheses.

\subsection{Clustering by evolutionary signatures reveals limited patterns of conserved features}
Given our observation of widespread signals of feature conservation in \textit{Drosophila} IDRs, we next sought to investigate if there are IDRs with similar patterns of conserved features and common functions. Accordingly, we treated the set of likelihoods ratios for each region as its ``evolutionary signature'' and clustered them as in as in Zarin \textit{et al.}~\cite{Zarin2019}. The resulting heatmap reveals patterns of similar signatures interspersed among a high background of noise (Fig.~\ref{fig:hierarchy}). However, we identified at least 39 clusters with strong and consistent patterns of constraint. To gain greater insight into their functional properties, we performed term enrichment analyses on the sets of proteins with regions in each cluster. After correcting for multiple comparisons using a 5\% false discovery rate, only five clusters were significantly enriched for any terms, which are given in Table~\ref{table:cluster_terms}. Most enriched terms are associated with fewer than ten proteins in each cluster, though the enriched terms for cluster 12 are markedly stronger and consistently associated with protein localization. We also observed that while many clusters are defined by a single strongly conserved feature, three (clusters 15, 23, and 24) have multiple features with signals of constraint. To determine if these more highly constrained regions are associated with a common underlying pattern of features, we compared their evolutionary signatures to normalized values of their inferred optima (Fig.~\ref{fig:clusters}B-D). The heatmaps reveal that while the regions in these clusters have common patterns of constraint, the optimal values of their features can vary considerably. Likewise, cluster 12's strongly constrained feature, repeat\_K, is similarly variable (Fig.~\ref{fig:clusters}A). Interestingly, this feature is consistent with its enrichment in annotations related to protein localization, as mitochondrial or nuclear targeting signals are characterized by a net positive charge or clusters of basic residues, respectively~\cite{Moulin2019, Lu2021}. However, the relative depletion of lysine repeats in some regions implies they are constrained against the formation of targeting signals. This in turn suggests that IDRs may in general have complex relationships between their evolutionary constraints and functions where constraints can both preserve beneficial functions as well as prevent the acquisition of deleterious ones.

\section{Discussion}
\subsection{IDRs have distinct patterns of sequence and feature evolution}
In this study, we applied several phylogenetic models to IDRs to interrogate the evolution of their sequences and molecular features. Most significantly, through a comparison of two models of continuous trait evolution we demonstrate evidence of widespread constraint in IDRs within the \textit{Drosophila} proteome. Furthermore, by quantitatively ranking the importance of various molecular features, these results can generate hypotheses to guide the functional dissection of IDRs in specific proteins. More broadly, though, it suggests that constraint of distributed features is a mechanism of IDR evolution common to multiple biological systems. Using evolutionary signatures derived from these models, we also attempted to identify clusters of IDRs with shared patterns of constraint and common functions. However, in contrast to the previous study in yeast, we find limited evidence of IDR clusters with specific biological functions. This may indicate a more complex regulatory environment in multicellular organisms which lacks a simple mapping between molecular features and functions. For example, as mentioned in the analysis of cluster 12, some constraints may correspond to the absence of a property or function. As annotations are typically framed as positive statements about function, enrichment analyses have less power to detect this type of negative relation. Alternatively, conserved features may correspond to classes of conformational ensembles or modalities of protein-protein interactions which are not necessarily associated with specific functions or processes. Technical differences between the two studies may also explain the discrepancy. In particular, the Brownian motion and Ornstein-Uhlenbeck processes are highly generic models of trait evolution and may have less power to detect signals of constrained evolution than the simulation approach used in the previous study. The lack of region-specificity in protein annotations further compounds these issues, as the spurious association of IDRs with annotations derived from other regions introduces noise to the enrichment analyses. However, these and other methodological limitations are discussed in greater detail in a subsequent section.

@@

\subsection{Future evolutionary analyses of IDRs will require a multimodal approach}
As discussed by Zarin \textit{et al.}, the interpretation of evolutionary signatures is complicated by several methodological limitations~\cite{Zarin2019}. For example, because IDRs are identified as contiguous segments of high predicted disorder, their boundaries are defined by adjacent structural elements. This approach can therefore split an IDR that is a single evolutionary or functional unit if it contains a semi-disordered module that scores below the threshold. Conversely, it can also merge two distinct IDRs if they are not separated by at least one folded domain. These issues can have significant effects on the accuracy of our statistical analysis, as mismatches between the inferred boundaries of an IDR and its true evolutionary or functional divisions can introduce noise or even spurious signals of constraint. However, more targeted applications of this framework where investigators manually choose the boundaries for specific IDRs can ensure the results are directly related to the regions of interest.

@@

However, the comparative phylogenetics methodology applied here also has limitations. As many features have strict boundaries or cannot vary continuously, they violate one or more of the underlying assumptions of the BM and OU models. Fortunately, for many features these inconsistencies likely do not seriously compromise the analysis. For example, though compositional features like fraction\_S are mathematically restricted to the interval between zero and one, they are likely constrained by much narrower selective regimes, and within these regimes, their behavior is effectively described by an OU model. For other features, however, the deviations are more consequential. For example, net\_charge and the motif features can only assume integer and non-negative values, respectively, which imposes significant restrictions on their allowed increments that are not reflected in the BM and OU models. Instead, more appropriate models for count data are birth-death processes, which are Markov chains defined on the natural numbers. However, though linear birth-death processes are well-studied and widely applied in biology~\cite{Crawford2011}, to our knowledge there are no simple parameterizations which describe a mean-reverting behavior analogous to the OU model. Thus, further theoretical developments are needed to apply birth-death processes as a model of stabilizing selection in studies of IDR evolution.

@@

\subsection{Estimation of Brownian motion and Ornstein-Uhlenbeck parameters}
Brownian motion (BM) model parameters were calculated with two methods. The first used Felsenstein's contrasts algorithm to efficiently calculate roots and contrasts for the disorder scores and features of each region~\cite{Felsenstein1973, Felsenstein1985}. Rates were calculated as the mean of the squares of the contrasts. Though these values are unbiased, they are not maximum likelihood estimates and are inappropriate for use with likelihood ratio calculations. Thus, they were used for analyses involving only the BM model. For comparison with the Ornstein-Uhlenbeck (OU) model, the BM parameters were calculated by maximizing the likelihood. The OU model parameters were also calculated via maximum likelihood estimation as described in Butler \textit{et al.}~\cite{Butler2004}. To ensure parameter identifiability and estimate consistency, the root was treated as a random variable~\cite{Ho2013, Ho2014}. Thus, the covariance matrix, $V$, was parameterized as $V_{ij} = e^{\alpha d_{ij}}$ where $d_{ij}$ is the tree distance between tips $i$ and $j$, and $\alpha$ is the selection strength~\cite{Ho2013}. Both the BM and OU models used the reference species tree to parameterize the branch lengths.

@@

\section{Acknowledgements}
We thank the past and present members of the Eisen lab for helpful discussions and their assistance in revising the manuscript.

>>>>

However, no known subsequent studies have determined if similar patterns of conservation are found in the IDRs of other systems. As another foundational model organism with abundant genomic information across many evolutionary lineages~\cite{Yang2018, Miller2018, Kim2021, Gramates2022}, the fruit fly, \textit{Drosophila melanogaster}, is a natural choice for subsequent investigation. Furthermore, given its complex multicellular development process and shared signaling pathways with humans, the findings of such a study would significantly advance our understanding of the role of IDRs in gene regulation as well as human health and disease. The concordance of these results with the previously identified IDR clusters would also have profound implications for the broader mechanisms of IDR evolution. For example, the absence of global patterns of evolutionary signatures across IDRs in \textit{Drosophila} would suggest they are property of IDRs which is unique to yeast. In contrast, the identification of clusters similar to those in yeast would support the existence of a taxonomy of IDRs which is conserved across the tree of life. The latter result would represent a significant step towards the creation of resources for the classification of IDRs analogous to those for folded domains such as Pfam~\cite{Mistry2020}, CATH~\cite{Sillitoe2020}, or SCOP~\cite{Andreeva2013, Andreeva2019}.

@@

\section{Results}
\subsection{IDRs are shorter and more divergent than non-disordered regions}
We identified regions with high levels of inferred intrinsic disorder in over 8,500 high quality alignments of single copy orthologs from 33 species in the \textit{Drosophila} genus using the disorder predictor AUCPreD~\cite{Wang2016}. To highlight the unique features of IDR evolution in subsequent analyses, we also extracted a complementary set of regions with low levels of inferred disorder. Both sets were filtered on several criteria, including the lengths of their sequences and their phylogenetic diversity, which yielded 11,445 and 14,927 regions, respectively, from 8,466 unique alignments. In the subsequent discussion, we refer to these sets as the ``disorder'' and ``order'' regions, respectively. To investigate the differences in basic sequence statistics between the two region sets, we first generated histograms from the average length of each region (Fig.~\ref{sfig:regions}). Although both distributions span several orders of magnitude, the order regions are generally longer than the disorder regions, with means of 105 and 245 residues, respectively. We then quantified the sequence divergence in each region by fitting phylogenetic trees to the alignments using amino acid and indel substitution models. The average rates of substitution are significantly larger in the disorder regions, demonstrating that while both sets contain conserved and divergent regions, IDRs are enriched in more rapidly evolving sequences (Fig.~\ref{sfig:regions}). We also searched the database of Pfam domains against the full-length \textit{D. melanogaster} proteins in these alignments and measured their overlap with the disorder and order regions. The results show the disorder set has a clear enrichment in regions with no or only small amounts of overlap relative to the order set, indicating that, compared to structured domains, IDRs are especially resistant to homology-based methods of functional annotation (Fig.~\ref{sfig:pfam}).

@@

\subsection{Intrinsic disorder is poorly conserved in some proteins}
Though the substitution models reveal specific patterns of evolution at the level of individual residues, the large amounts of sequence divergence between many orthologous IDRs implies their evolution is not well-described by fine-scale models of residue substitution. Given the growing evidence that IDRs are constrained to conserve distributed properties, we instead turned towards characterizing their evolution in terms of 82 disorder-associated ``molecular features'' obtained from the previous study of IDRs in the yeast proteome. However, before conducting an in-depth analysis of these features, we examined the disorder score traces in greater detail and were struck by the significant variability between species. For each residue in the input sequence, AUCPreD returns a score between 0 and 1 where higher values indicate higher confidence in a prediction of intrinsic disorder. In some alignments, the scores vary by nearly this entire range at a given position even when there is a relatively high level of sequence identity (Fig.~\ref{fig:scores}A).

@@

\subsection{IDRs have three axes of unconstrained variation}
Having calculated the features associated with the sequence segments composing each region in our data set, we then sought to determine if their distributions contained any global structure which would enable us to identify classes with distinct biophysical or functional properties. These distributions are generated by a complex underlying evolutionary process which reflects the combined effects of selection, drift, and mutation. However, to leverage a statistical framework to infer the properties of this process, we fit Brownian motion (BM) models to the features calculated from the segments in each region. BM is a simple model of evolution where continuous traits change through a series of small, undirected steps. Thus, the traits accumulate variation at a constant rate over time but do not on average deviate from their original values. BM models are therefore specified by two parameters: a rate ($\sigma^2_{OU}$), which describes the speed at which trait variation accumulates, and a root ($\mu_{BM}$), which describes the ancestral trait value.

We then applied principal components analysis (PCA) to visualize the major axes of variation of the root and rate parameters for each feature and region. A difficulty with a direct analysis of the parameter estimates, however, is the sensitivity of PCA to differences in scaling between variables, and some features have dramatically different intrinsic scales. For example, many compositional features, like fraction\_S, are restricted to the interval [0, 1], whereas SCD is unbounded and can vary from negative to positive infinity. As a result, SCD is responsible for a significant fraction of the overall variance in both parameter distributions (Fig.~\ref{sfig:root_variance}-\ref{sfig:rate_variance}). Therefore, we first normalized the parameters associated with each feature by transforming them into $z$-scores relative to their proteome-wide distributions.

The first two principal components of the root distributions show little overall structure, though there is a slight enrichment of regions along two axes that correlate with acidic and polar features, respectively (Fig.~\ref{sfig:root}). Likewise, the projections of the rates onto the first two principal components are largely distributed along the first (Fig.~\ref{fig:rate}A). This observation and the variable amounts of sequence divergence in the regions led us to suspect the first principal component was a measure of the overall rate of sequence evolution. Plotting the first principal component against the sum of the average amino acid and indel rates as measured by substitution models revealed a strong association (Fig.~\ref{fig:rate}B). We then projected the rates along second and third principal components to determine if these higher order components contained any additional structure. The resulting distribution is roughly triangular and contains three major axes of variation, corresponding to rapid changes in the regions' proportions of acidic, glutamine, and glycine residues (Fig.~\ref{fig:rate}C-D). Inspection of regions selected along these axes confirmed the high rates of evolution of these features (Fig.~\ref{fig:rate}E-G). Furthermore, we observe a similar distribution when the rates of the order regions were projected along their second and third principal components, which suggests a lack of constraint along these axes is a general property of rapidly evolving proteins (Fig.~\ref{sfig:order}).

\subsection{A model of constrained evolution can identify signals of conservation}
Though the BM process permits the inference of the rates of feature evolution after accounting for the phylogenetic relationships between species, it does not directly test for their conservation. In fact, under the BM model, trait variation is unconstrained and will increase without bound over time. Instead, evidence of conservation requires comparison to a model where trait variation is constrained. A common choice for modeling the effect of selection on the evolution of a continuous trait is the Ornstein-Uhlenbeck (OU) model. The OU model is similar to the BM model where a trait accumulates variation through a series of small, undirected steps. However, it differs in that the trait is also attracted towards an optimal value where the attraction is proportional to the trait's distance from this value. Under an additional assumption of stationarity that ensures parameter identifiability and estimate consistency, the OU model is accordingly specified with three total parameters: the optimal value ($\mu_{OU}$), the fluctuation magnitude ($\sigma^2_{OU}$), and the selection strength ($\alpha$)~\cite{Ho2013, Ho2014}. While the first two parameters are analogous to the root and rate parameters in the BM model, respectively, the selection strength has no equivalent.

@@

We then quantified the type I and II error rates under a hypothesis testing framework. We first calculated the type I error rate as a function of the critical value of the log likelihood ratio for each value of the BM rate separately~(Fig.~\ref{fig:simulation}D). As the overlapping curves indicate the ratios are independent of the true value, we merged the simulations and calculated empirical critical values for 5\% (2.58) and 1\% (4.20) type I error rates~(Fig.~\ref{fig:simulation}E). Using the 1\% critical value, we calculated the type II error rate as a function of the OU parameter values~(Fig.~\ref{fig:simulation}F). The results show a strong dependence on the log ratio of $\alpha$ to $\sigma^2_{OU}$, with the error rate sharply declining to zero when $\log_{10}(\frac{\alpha}{\sigma^2_{OU}}) > 1$. In summary, these results demonstrate that while the parameter estimates of the OU model can have significant bias, when selection is strong relative to the fluctuation magnitude, the log likelihood ratio of the two models can reliably signal against a null hypothesis of Brownian motion.

\subsection{Signals of feature conservation are widespread in IDRs}
Having evaluated the statistical properties of the hypothesis testing framework, we next sought to investigate the distribution of significant features across all regions. However, to ensure we would not identify signals of conservation simply as a result of high levels of sequence identity, we restricted our subsequent analyses to disorder regions with a minimum amount of divergence as measured by substitution models (Fig.~\ref{sfig:hierarchy_histogram}). We first quantified the fraction of regions with significant log likelihood ratios using the empirical critical values for 1 and 5\% type I error rates, finding that all non-motif features show significance rates that greatly exceed their nominal error rates (Fig.~\ref{fig:conserved}A). The motif features, though, are more variable, with some reaching significance rates similar to those of the non-motif features, whereas many others have rates at or below the nominal values. Generally, the shorter and more degenerate motifs associated with modification sites have higher rates of significance, which may reflect stronger constraints on motifs which are more likely to arise by random mutations. We next calculated the number of significant features in each region using the same critical values. Both yielded highly skewed distributions, indicating the observation of a significant feature increases the probability of observing another significant feature (Fig.~\ref{fig:conserved}B). However, as many features have overlapping definitions, this dependence is expected. To correct for this effect, we randomly shuffled the log likelihood ratios between regions for each feature independently and calculated the overall significance rate across all features. For both critical values, the empirical significance rate exceeds the nominal error rate. Thus, constrained feature evolution is widespread in IDRs, with each region containing an average of seven significant features under the 1\% type I error rate.

To illustrate the utility of this method for generating hypotheses to guide the functional dissection of specific proteins, we next compared two regions with elevated values of fraction\_Q and repeat\_QN but different signals of conservation for each (Fig.~\ref{fig:alignment}A-B). Though the alignments of both regions are enriched in glutamine and asparagine residues (Fig.~\ref{fig:alignment}C,E), the first has significant signals of conservation for both features (Fig.~\ref{fig:alignment}D), whereas the second does not (Fig.~\ref{fig:alignment}F). Notably, the first region is from an orthologous group that contains a sequence from the \textit{D. melanogaster} gene {ewg}, which is a transcription factor. As polyglutamine stretches can modulate the activity of transcription factor activation domains~\cite{Mitchell1989, Gerber1994}, this observation is consistent with the maintenance of an optimal level of transcriptional activation mediated by constraints on glutamine-associated features. In contrast, the low log likelihood ratios of the second region, whose \textit{D. melanogaster} gene is unannotated, are suggestive of unconstrained glutamine repeat expansion and contraction. Interestingly, this region has highly significant log likelihood ratios for several features associated with glycine or phenylalanine repeats despite having few of these residues. However, careful inspection of the alignment reveals several positions where such repeats appear in only one or two sequences, \textit{e.g.} columns 153 and 208. While these strong signals of conservation may represent artifacts of applying a continuous model of trait evolution to features derived from discrete data, they could also indicate true constraints against the presence of flexible or hydrophobic patches in those sequence segments. Taken together, these results highlight this method's ability to identify biochemical features with signals of constrained evolution while also emphasizing the importance of combining this information with a protein's biological context to jointly suggest testable hypotheses of function.

\subsection{Clustering by evolutionary signatures reveals limited patterns of conserved features}
Given our observation of widespread signals of feature conservation in \textit{Drosophila} IDRs, we next sought to investigate if there are IDRs with similar patterns of conserved features and common functions. Accordingly, we treated the set of log likelihoods ratios for each region as its ``evolutionary signature'' and clustered them as in Zarin \textit{et al.}~\cite{Zarin2019}. The resulting heatmap reveals patterns of similar signatures interspersed among a high background of noise, and we identified at least 39 clusters with strong and consistent patterns of constraint (Fig.~\ref{fig:hierarchy}). To gain greater insight into their functional properties, we performed term enrichment analyses using the annotations associated with each region's protein. After correcting for multiple comparisons using a 5\% false discovery rate, only five clusters were significantly enriched for any terms, which are given in Table~\ref{table:cluster_terms}. Most enriched terms are associated with fewer than ten proteins in each cluster, though the enriched terms for cluster 12 are markedly stronger and consistently associated with protein localization. We also observed that while many clusters are defined by a single strongly conserved feature, three (clusters 15, 23, and 24) have multiple features with signals of constraint. To determine if these more highly constrained regions are associated with a common underlying pattern of features, we compared their evolutionary signatures to normalized values of their inferred optima (Fig.~\ref{fig:clusters}B-D). The heatmaps reveal that while the regions in these clusters have common patterns of constraint, the optimal values of their features can vary considerably. Likewise, cluster 12's strongly constrained feature, repeat\_K, is similarly variable (Fig.~\ref{fig:clusters}A). Interestingly, this feature is consistent with its enrichment in annotations related to protein localization, as mitochondrial or nuclear targeting signals are characterized by a net positive charge or clusters of basic residues, respectively~\cite{Moulin2019, Lu2021}. However, the relative depletion of lysine repeats in some regions implies they are constrained against the formation of such targeting signals. This interpretation in turn suggests that IDRs may in general have complex relationships between their evolutionary constraints and functions where constraints can both preserve beneficial functions as well as prevent the acquisition of deleterious ones.

\section{Discussion}
\subsection{IDRs have distinct patterns of sequence and feature evolution}
In this study, we applied several phylogenetic models to IDRs to interrogate the evolution of their sequences and molecular features. Most significantly, through a comparison of two models of continuous trait evolution we demonstrate evidence of widespread constraint in IDRs within the \textit{Drosophila} proteome. Furthermore, by quantitatively ranking the importance of various molecular features, these results can generate hypotheses to guide the functional dissection of IDRs in specific proteins. More broadly, though, it suggests that constraint of distributed features is a mechanism of IDR evolution common to multiple biological systems. Using evolutionary signatures derived from these models, we also attempted to identify clusters of IDRs with shared patterns of constraint and common functions. However, in contrast to the previous study in yeast, we find limited evidence of IDR clusters with specific biological functions. This may indicate a more complex regulatory environment in multicellular organisms which lacks a simple mapping between molecular features and functions. For example, as mentioned in the analysis of cluster 12, some constraints may correspond to the absence of a property or function. As annotations are typically framed as positive statements about function, enrichment analyses have less power to detect this type of negative relationship between a feature and its function. Alternatively, conserved features may correspond to classes of conformational ensembles or modes of molecular interaction which are not necessarily associated with specific functions. Technical differences between the two studies may also explain the discrepancy. In particular, the BM and OU processes are highly generic models of trait evolution and may have less power to detect signals of constrained evolution than the simulation approach used in the previous study. The lack of region-specificity in annotations further compounds these issues, as the spurious association of IDRs with annotations derived from other regions introduces noise to the enrichment analyses. However, these and other methodological limitations are discussed in greater detail in a subsequent section.

@@

\subsection{Future evolutionary analyses of IDRs will require a multimodal approach}
As discussed by Zarin \textit{et al.}, the interpretation of evolutionary signatures is complicated by several methodological limitations~\cite{Zarin2019}. For example, because IDRs are identified as contiguous segments of high predicted disorder, their boundaries are defined by adjacent structural elements. This approach can therefore split an IDR that is a single evolutionary or functional unit if it contains a semi-disordered module that scores below the threshold. Conversely, it can also merge two distinct IDRs if they are not separated by at least one folded domain. These issues can have significant effects on the accuracy of our analyses, as mismatches between the inferred boundaries of an IDR and its true evolutionary or functional divisions can introduce noise or even spurious signals of constraint. However, more targeted applications of this framework where investigators manually choose the boundaries for specific IDRs can ensure the results are directly related to the regions of interest.

@@

However, the comparative phylogenetics methodology applied here also has limitations. As many features have strict boundaries or cannot vary continuously, they violate one or more of the underlying assumptions of the BM and OU models. Fortunately, for many features these inconsistencies likely do not seriously compromise the analysis. For example, though compositional features like fraction\_S are mathematically restricted to the interval between zero and one, they are likely constrained by much narrower selective regimes, and within these regimes, their behavior is effectively described by an OU model. For other features, however, the deviations are more consequential. For example, net\_charge and the motif features can only assume integer and natural number values, respectively, which imposes significant restrictions on their allowed increments that are not reflected in the BM and OU models. Instead, more appropriate models for count data are birth-death processes, which are Markov chains defined on the natural numbers. However, though linear birth-death processes are well-studied and widely applied in biology~\cite{Crawford2011}, to our knowledge there are no simple parameterizations which describe a mean-reverting behavior analogous to the OU model. Thus, further theoretical developments are needed to apply birth-death processes as a model of stabilizing selection in studies of IDR evolution.

@@

\subsection{Estimation of Brownian motion and Ornstein-Uhlenbeck parameters}
Brownian motion (BM) model parameters were calculated with two methods. The first used Felsenstein's contrasts algorithm to efficiently calculate roots and contrasts for the disorder scores and features of each region~\cite{Felsenstein1973, Felsenstein1985}. Rates were calculated as the mean of the squares of the contrasts. Though these values are unbiased, they are not maximum likelihood estimates and are inappropriate for use with log likelihood ratio calculations. Thus, they were used for analyses involving only the BM model. For comparison with the Ornstein-Uhlenbeck (OU) model, the BM parameters were calculated by maximizing the likelihood. The OU model parameters were also calculated via maximum likelihood estimation as described in Butler \textit{et al.}~\cite{Butler2004}. To ensure parameter identifiability and estimate consistency, the root was treated as a random variable~\cite{Ho2013, Ho2014}. Thus, the covariance matrix, $V$, was parameterized as $V_{ij} = e^{\alpha d_{ij}}$ where $d_{ij}$ is the tree distance between tips $i$ and $j$, and $\alpha$ is the selection strength~\cite{Ho2013}. Both the BM and OU models used the reference species tree to parameterize the branch lengths.

@@

\section{Acknowledgements}
We thank the past and present members of the Eisen lab and Max Staller for helpful discussions and their assistance in revising the manuscript.
